%
%  mindsight
%
%  Created by Amiel Martin on 2010-01-18.
%  Copyright (c) 2010 Amiel Martin. All rights reserved.
%
\documentclass[]{article}

% Use utf-8 encoding for foreign characters
\usepackage[utf8]{inputenc}

% Setup for fullpage use
\usepackage{fullpage}

\usepackage{setspace} 
\doublespacing

% Uncomment some of the following if you use the features
%
% Running Headers and footers
\usepackage{fancyhdr}

% Package for including code in the document
% \usepackage{listings}


\title{ On Attachment \\
	\large some ramblings on Attachment Theory }
\author{ Amiel Martin }

\date{2010-02-18}

\begin{document}
	\maketitle
	
	\subsection*{Introduction} % (fold)
	\label{sub:introduction}
	
	I feel that a lot of what we've been learning in this class is stuff that everyone should know.
	When I tell somebody about what I'm learning in the class, and I get to attachment theory, which I feel is at the heart of most of this, I explain the four attachment styles, but I realize that I have not been representing it very well.
	I blur the lines of attachment styles in children and with adults; I don't offer evidence for how important this stuff is; I just offer resources to learn more about attachment theory because I'm not doing it justice.
	We've been exploring attachment theory and related studies in detail and I'd like to step back a little bit and explore how the basics and the importance of attachment theory can be taught to anyone in a short amount of time.
	In order to do this, I'll probably have to explore attachment theory in more detail and try to pick out the good bits, the interesting bits, the powerful bits, and a simple overview.
	What I'm doing here is revisiting attachment theory, the basic premise of this class as I understand it, in order to be able to teach it more succinctly to other people.
	
	So... where to start.
	I kind of like introducing some of this stuff with the concept that science is now teaching us what a lot of people already knew.
	Take for example, the relatively recent view in science that relationships have an impact on mental and physical health.
	Cacioppo and Patrick wrote that our social nature is part of evolutionary development\footnote{in \emph{selfish genes social animals}}.
	Being socially excluded activates the same circuits in our brain as physical pain.
	In other words, being socially excluded hurts. Well DUH!
	
	% subsection introduction (end)
	
	
	A really important field of study primarily from the last thirty years, attachment theory, was originally formulated by John Bowlby.
	His main point is that there is ``a strong causal relationship between and individual's experiences with his parents and his later capacity to make affectional bonds,''\footnote{from \emph{The Making \& Breaking of Affectional Bonds}} and that the ways in which relationships later in life are developed can be attributed to the ways that parents perform their roles.
	Bowlby introduced how infants form bonds with their caregivers, and that they either have secure or insecure attachment to their caregiver.
	Securely attached children tend to grow to be self-reliant, trusting, cooperative, and helpful to others.
	Most importantly, securely attached children feel that they are able to help other people and that they themselves are worthy of help should difficulties arise.
	Bowlby explains that secure attachment is raised when the parents have an understanding of the child's attachment behaviors and a willingness to meet them, as well as ``recognition that one of the commonest sources of a child's anger is the frustration of his desire for love and care.''
	Bowlby writes about two opposite patterns of attachment: anxious and compulsive self-reliance. I'll come back to these.
	
	 Around this same time Harry Harlow was working on his famous surrogate mother experiment\footnote{ Some footage of Harry Harlow's experiments can be found on youtube: http://www.youtube.com/watch?v=fLrBrk9DXVk }.
	With extreme brevity: Harlow removed baby rhesus monkeys from their mothers and instead offered various combinations of terrycloth and wire ``mothers.'' \footnote{ An overview of the study with links to more resources can be found on wikipedia at http://en.wikipedia.org/wiki/Harry\_Harlow\#Surrogate\_mother\_experiment }
	He found that even when the wire mother was source of food, monkeys preferred cuddling with the terrycloth mother, and even found it to be a source of comfort.
	Monkeys that were raised with a terrycloth mother would seek proximity to it when introduced with a frightening stimulus, and would soon gain the confidence to explore.
	Monkeys that were raised with only a wire mother were not able to be comforted enough to explore.
	
	Another important study in the development of attachment theory was the ``Strange Situation,'' developed by Mary Ainsworth.
	In the ``Strange Situation,'' infants and toddlers were brought into a room with their parent and a stranger.
	Observations were made on how the child acted when the parent left with or without the stranger, and how the child acted when they returned. 
	Because research is research, and even though you can't put people in boxes, you still have to define categories, they defined four types of attachment: Secure, Avoidant, Resistant or Ambivalent, and Disorganized.
	There is certainly a lack of standardization of terms in the field of attachment theory; what Ainsworth calls Avoidant is what Bowlby calls compulsive self-reliance, and Resistant or Ambivalent is what Bowlby calls Anxious. For a visual overview of the four attachment styles see Table~\ref{table:attachment_styles}.\footnote{Quoted from Wikipedia: \small{http://en.wikipedia.org/wiki/Attachment\_theory\#Attachment\_patterns}}

	\begin{table}
		\caption{ Attachment Styles }
		\label{table:attachment_styles}
		
		\begin{tabular}{ | p{1in} | p{3.3in} | p{1.7in} | }
			\hline
			\textbf{Attachment pattern} & \textbf{Child} & \textbf{Caregiver} \\
			\hline
			\textbf{Secure} & Uses caregiver as a secure base for exploration. Protests caregiver's departure and seeks proximity and is comforted on return, returning to exploration. May be comforted by the stranger but shows clear preference for the caregiver. & Responds appropriately, promptly and consistently to needs. \\
			\hline
			\textbf{Avoidant} & Little affective sharing in play. Little or no distress on departure, little or no visible response to return, ignoring or turning away with no effort to maintain contact if picked up. Treats the stranger similarly to the caregiver. & Little or no response to distressed child. Discourages crying and encourages independence. \\
			\hline
			\textbf{Ambivalent / Resistant} & Unable to use caregiver as a secure base, seeking proximity before separation occurs. Distressed on separation with ambivalence, anger, reluctance to warm to caregiver and return to play on return. Preoccupied with caregiver's availability, seeking contact but resisting angrily when it is achieved. Not easily calmed by stranger. & Inconsistent between appropriate and neglectful responses. \\
			\hline
			\textbf{Disorganised} & Stereotypies on return such as freezing or rocking. Lack of coherent attachment strategy shown by contradictory, disoriented behaviours such as approaching but with the back turned. & Frightened or frightening behaviour, intrusiveness, withdrawal, negativity, role confusion, affective communication errors and maltreatment. \\
			\hline
		\end{tabular}
	\end{table}
	
	
	Here are a few statements that from what I understand of attachment styles. They might be oversimplifying and therefor possibly misleading.
	I've found that I had a hard time getting this out because I have a hard time explaining them without tact.
	Secure attachment allows a person to feel worthy of being cared for and capable of giving care and help to others.
	Also, securely attached people have more ability to have empathy, and to articulate their past experiences.
	Studies have found that secure attachment is fostered by caregivers who are sensitive to the child's signals, or ``tuned in'' to their emotional state of mind. It has also been proposed that the ability to reflect on one's own as well as others state of mind is at the heart of many secure attachments.
	
	% to terse
	The Avoidant attachment style has a dominant left brain, and less access to autobiographical memories. 
	This lack of access to memories about feelings, hinders their understanding of other peoples feelings, empathy.
	The Avoidant attachment style is caused by parents who have less attunement and connection with their children.
	They have learned that they cannot find emotional support when they need it, so they are more likely to shut down feelings instead of seeking support.
	
	The stereotype of the ambivalent or anxious style of attachment is needy and clingy and doesn't feel that they are worthy of care. --- or, they don't feel like an individual without someone else.
	They are needy because love was inconsistent growing up.
	
	Both Avoidant and Ambivalent attachment styles are coping strategies.
	The Avoidant stategy is to block out feelings because nobody has helped you name them, understand them, and learn how to deal with them.
	The Ambivalent strategy is to be neurotic about care because it was available sometimes but not always when they needed it.
	The fourth attachment style is the outcome to what there is no coping strategy, when the source of care is the same source of fear. This is called the Disorganized attachment style, and is caused by parents who are frightening or act frightened.
	
	This is my basic understanding of Attachment theory.
	I think that my descriptions of each attachment style are a bit terse, but describe the general idea.
	Before I finish here, I would like to share one fact from a fascinating study by Marinus H. Van Ijzendoorn Abrahamsagi.
	In \emph{Cross-Cultural Patterns of Attachment}, she writes about studies of attachment theory in other cultures.
	In particular, the Gusii of Kenya, who have multiple caregivers, offer some interesting insight to attachment theory.
	Gusii mothers give physical care to children, while child caregivers offer social and playful interactions.
	What they found was that Gusii children that were securely attached to their mothers had good physical health, and that Gusii children that were securely attached to their child caregivers were more intelligent.
	
	So, there you go. That was my attempt to clarify attachment theory for myself.
	Did it clarify anything for you?
	I hope that I at least peaked your interest enough to go read more because this stuff is so important, especially if you ever plan to raise or help raise a child.
	I recommend the following resources: Bowlby's writings on attachment offer the original ideas\footnote{ Bowlby. \emph{The Making and Breaking of Affectional Bonds}}; Daniel Siegel's chapter on attachment from the Developing Mind explains the attachment styles and gives lots of details, if you only read one this one is probably the one to read\footnote{ Daniel Siegel. Chapter 3 in \emph{The Developing Mind}}; Mikulincer and Shaver\footnote{ Chapter 2 in \emph{Attachment in Adulthood}}; and Daniel Siegel's Mindsight book has a really great chapter which explains the basics of attachment theory and gives lots of good stories\footnote{ Daniel Siegel. Chapter 9 ``Making Sense Of Our Lives'' from \emph{Mindsight} }.


	% Immune defenses are buffered by love and support.\footnote{Dean Ornish, \emph{Love \& Survival}}
	% Biological basis for affiliation, it promotes health, and the absence of relationships creates illness.
	% 
	%	


	% what works in a speech
	% 
	% include a story about yourself
	% formula for speech:
	% 	intro:
	% 		core message
	% 	story in your life or someone else where that message became clear to you
	% 	info
	% 	wrap up:
	% 		consider this! (that you can make a difference)
	% 	call to action
	% 	
	% ask a question
	% 



\end{document}
