
%
%  untitled
%
%  Created by Amiel Martin on 2010-03-11.
%  Copyright (c) 2010 __MyCompanyName__. All rights reserved.
%
\documentclass[]{article}

% Use utf-8 encoding for foreign characters
\usepackage[utf8]{inputenc}

% Setup for fullpage use
\usepackage{fullpage}

% Uncomment some of the following if you use the features
%
% Running Headers and footers
\usepackage{fancyhdr}

% Package for including code in the document
% \usepackage{listings}


\title{Experiences for Mindsight Presentation}
\author{ Amiel Martin }

\date{2010-03-11}

\begin{document}
	\maketitle
	
	\begin{enumerate}
		\item Introduce the presentation
			
			``I'm working on this workshop, and because of the time constraint, I'd like to just show you the experiential parts...''
			
		\item Science is teaching us what we already know...
		
			Social rejection and physical pain share the same neural pathways, in other words Social rejection hurts. Well, DUH!
			
			Attachment theory teaches us to raise our children with love and care. Well, DUH!
		
		\item Neuroplasticity
		
			``One of the key practical lessons of modern neuroscience is that the pawer to direct our attention has with in it the power to shape our brain's firing patterns, as well as the power to shape the architecture of the brain itself.''
			
			\begin{itemize}
				\item Explain that our brains are plastic, and can change.
				\item Analogy of the trail in the field.
				
					Every time you walk it, you dig a path, which can get deeper and become a trench. These are your habits, and you can get out of any trench, but the deeper it gets, the more work it is to get out. You might have to dig a staircase in the side, but you can do it.
					
				\item ``No Guns'' skill practice
				
					\begin{itemize}
						\item Alternate finger and thumb for each hand
						\item Do both synchronous
						
							``Notice how easy this is, and that there is nothing muscular keeping you from doing it.''
							
						\item Alternating (the difficult version), and ways to practice it.
						
							``By practicing this, you are building new neural circuits. And when you get it, it becomes easy.''
							
					\end{itemize}
					
			\end{itemize}
			
		\item Reactivity and Receptivity
		
			\begin{itemize}
				\item Define them
				\item Tell some stories about reactive times (maybe have people tell their own stories, not sure if this is a good idea)
				\item Hand model of the brain, Functions of the Middle Prefrontal Cortex, and how we can ``flip our lid''

				Bodily Regulation, Attuned Communication, Emotional Balance, Fear Suppression, Response Flexibility, Insight, Empathy, Morality, and Intuition.
				
				\item Reactivity experience
				
					``Sit comfortably, close your eyes, and just notice what sensations arise.''
				
					NO, NO, NO, NO, NO, NO, NO...  yes, yes, yes, yes, yes, yes
				
			\end{itemize}
			
		\item Mindfulness exercise
		
			\begin{itemize}

				\item ``Here is an exercise you can do to help become more aware of your states of mind. There are many others, such as yoga...''
				\item Studies have shown that people with ADD improved more and with lasting effects by meditation than medication.
				\item * Read meditation from Daniel Siegel's \emph{Mindsight} book
				\item ``How did that feel?''
				\item Encourage people to do it at home, ``brain brushing?...''

			\end{itemize}
			
	\end{enumerate}

\end{document}


Brain Plasticity
	juggling
	visualization
